---
title: "Sample Size Calculation for Clinical Trials with Multiple Primary Endpoints"
author: "Srinath Kolampally and Matthias Kohl"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteEngine{rmarkdown::render}
  %\VignetteIndexEntry{MPE}
  %\VignetteEncoding{UTF-8}
---


##                    1. Introduction  


The calculation of sample size or power is a difficult task when multiple primary 
endpoints (MPE) are considered, which means when there is more than one primary endpoint. 
In general, these MPE are categorized into multiple co-primary and multiple primary 
endpoints. A trial with multiple co-primary endpoints is successful if there is
a significant improvement for **ALL THE ENDPOINTS** whereas in case of multiple 
primary endpoints a success in **AT LEAST ONE ENDPOINT** is sufficient. The 
package **mpe** can be used to calculate sample size and power for MPE.

First, we give some details about the methods and formulas. Next, we illustrate 
how to apply our functions in simple examples. Finally, we consider Alzheimer's Disease (AD) as an application.


## 2. Multiple Co-primary Endpoints

In case of co-primary endpoints, the trial is designed to evaluate ALL THE ENDPOINTS. 
For example, in a clinical trial evaluating some treatments in 
patients affected by irritable bowel syndrome (IBS) the FDA recommends to evaluate 
ALL endpoints for assessing the IBS signs and symptoms. The co-primary endpoints are: 

1. pain intensity and stool frequency of IBS with constipation 
2. pain intensity and stool consistency of IBS with diarrhea (FDA 2012).

### 2.1. Known Covariance

In this section, we briefly state the relevant formulas from Chapter 2 of the book 
by Sozu et al. [12], where we assume the covariance as known. 
Let $K$ be the number of co-primary endpoints and $Y_{C_k}$ and $Y_{T_k}$ ($k=1,\ldots,K$) 
be the results for the control (C) and test (T) group in a randomized superiority 
clinical trial. $Y_C=(Y_{C_1},\ldots,Y_{C_K})$ and $Y_T=(Y_{C_1},\ldots,Y_{C_K})$ 
are assumed to follow a $K$-variate normal distribution with mean vector 
$µ_C = (µ_{C_1},\ldots, µ_{C_K})$ and $µ_T = (µ_{T_1},\ldots,µ_{T_K})$
and (known!) covariance matrix $∑$; i.e., $Y_C \sim N_k(µ_C,∑)$ and $Y_T\sim N_k(µ_T,∑)$.
   
where 
     
$$\mathbf{\Sigma} = \left[\begin{array}
{ccc}
\sigma_1^2 & ... & {\rho^1_K}{\sigma_1}{\sigma_K}\\
 \vdots &  \vdots & \vdots   \\
{\rho^1_K}{\sigma_1}{\sigma_K}&  ... &  \sigma_K^2
\end{array}\right]$$
   
As we are considering a superiority trial, the superiority of the test intervention 
over control in terms of all the $K$ primary endpoints is given if and only if 
$µ_{T_k} - µ_{C_k} > 0$ for all $k = 1,2.....K$. In statistical testing, the 
null hypothesis $H_O$ is rejected if and only if all of the null hypotheses 
associated with each of the $K$ primary endpoints are rejected at a significance 
level of $\alpha$. The rejection region is an intersection of the involved $K$ 
co-primary endpoints and the statistical test used for the analysis is called 
intersection-union test [2]. That is, we consider the following hypotheses
   
   >$H_0 : µ_{T_k} - µ_{C_k} <= 0$, for at least one $k$,

   >$H_1 : µ_{T_k} - µ_{C_k} > 0$, for all $k$.
     
As we are assuming the covariance to be known, the hypotheses can be tested 
using a multivariate (intersection-union) z-test implemented in function 
`mpe.z.test`. The test statistic reads
 
 $$Z_k = \frac{\sqrt{n}({\overline{Y_{T_k}} -\overline{Y_{C_k}}})}{\sqrt{2}\sigma_k},
 \quad  k = 1,2,\ldots, K$$ 

where $n$ is the sample size *per group* (i.e., we assume a balanced design), 
and $\overline{Y_{T_k}}$ and $\overline{Y_{C_k}}$ are the respective sample means.
 
The overall power of this test is

$$1-\beta = P\left(\bigcap_{k=1}^K \{Z_k > z_\alpha\}\,\Big|\, H_1\right)$$  
                    
where $z_\alpha$ is the $(1-\alpha)$-quantile of the standard normal distribution and the overall power is referred to as conjunctive power [3, 10].

Next, we demonstrate how to calculate the sample size for a trial with two 
co-primary endpoints with known covariance.  
 
* K = 2, number of co-primary endpoints 
* delta = c(0.25, 0.4), expected effect size
* either Sigma or rho and SD must be specified
* Sigma = matrix(c(1, 0.8, 0.8, 1), nrow = 2), known covariance matrix
* rho = 0.8, known correlation 
* SD = c(1,1), known standard deviations
* sig.level = 0.025, significance level 
* power = 0.80, intended power
* n = ?

```{r}
library(mpe)
Sigma <- matrix(c(1, 0.8, 0.8, 1), nrow = 2)
power.known.var(K = 2, delta = c(0.25, 0.4), Sigma = Sigma, sig.level = 0.025, power = 0.8)
```

```{r}
## equivalent: known SDs and correlation rho
power.known.var(K = 2, delta = c(0.25, 0.4), SD = c(1,1), rho = 0.8, sig.level = 0.025, 
                power = 0.8)

```
We obtain a sample size of n = 252 which is the sample size per group; that is, 
the total sample size required is 2n = 504.

On the other hand, we can also calculate the power for a given sample size. In the above example with two multiple co-primary endpoints with known covariance we get. 
      
* K = 2, number of co-primary endpoints 
* delta = c(0.25, 0.4), expected effect size 
* Sigma = matrix(c(1, 0.8, 0.8, 1), nrow = 2), known covariance matrix
* rho = 0.8, known correlation 
* SD = c(1,1), known standard deviations
* sig.level = 0.025, significance level 
* n = 252, sample size per group
* power = ?

```{r}
power.known.var(K = 2, n = 252, delta = c(0.25,0.4), Sigma=Sigma, sig.level = 0.025)
```

```{r}
## equivalent: known SDs and correlation rho
power.known.var(K = 2, n = 252, delta = c(0.25, 0.4), SD = c(1,1), rho = 0.8, 
                sig.level = 0.025)

```
    
The computed power is $80.1\%$.
     

Furthermore, we can perform the corresponding multivariate intersection-union z-test
by applying our function `mpe.z.test`. Here we use simulated data for the 
demonstration.

```{r}
## effect size
delta <- c(0.25, 0.5)
## covariance matrix
Sigma <- matrix(c(1, 0.75, 0.75, 1), ncol = 2)
## sample size
n <- 50
## generate random data from multivariate normal distributions
X <- rmvnorm(n=n, mean = delta, sigma = Sigma)
Y <- rmvnorm(n=n, mean = rep(0, length(delta)), sigma = Sigma)
## perform multivariate z-test
mpe.z.test(X = X, Y = Y, Sigma = Sigma)
```


###2.2. Unknown Variance

In this section, we give the formulas from Chapter 2 of [12] where we assume the 
covariance as unknown. Let $D=(D_{1},\ldots,D_{K})^T$ with $D_{K}=\frac{\overline{Y_{T_k}} - \overline{Y_{C_k}}} {\sigma_{k}}$ be the effect sizes for the $K$ endpoints. Then, 
$\sqrt{kn}{D}$ is distributed as a k-variate normal distribution with mean vector 
$\sqrt{kn}{\delta}$ and covariance matrix $\rho_Z$; i.e., 
$\sqrt{kn}{D}\sim N_k(\sqrt{kn}{\delta},\rho_z)$. 

The hypotheses given above in Sec. 2.1 can be tested using a multivariate 
intersection-union t-test, which is implemented in function `mpe.t.test`. The 
test statistic is 
 
  $$T_k = \frac{\sqrt{n}({\overline{Y_{T_k}}-\overline{Y_{C_k}}})}{\sqrt{2}{s_k}},
  \quad  k = 1,2,\ldots, K$$
 
 
where $n$ is the sample size *per group* (i.e., we assume a balanced design), and 
$s_{k}$ is the estimate of the pooled standard deviation; 
$\overline{Y_{T_k}}$ and $\overline{Y_{C_k}}$ are the respective sample means.

 
The overall power reads

$$1-\beta = P\left(\bigcap_{k=1}^K \{T_k > t_{\alpha,{2n-2}\}},|\, H_1\right)$$ 


where $t_{\alpha,{2n-2}}$ is the $(1-\alpha)$-quantile of the t-distribution 
with $2n-2$ degrees of freedom.
  
  
Now, we demonstrate how to calculate the sample size for a trial with two 
co-primary endpoints with unknown covariance. Here, we follow three steps to 
determine the sample size.

+ Step 1: As we need starting values for our algorithm that computes the sample 
size in this case, we first act as if the covariance would be known and compute 
the sample size by applying our function `power.known.var`.

+ Step 2: The resulting value of `n` is considered as lower bound for the sample 
size in case of unknown covariance and is used as `min.n` in function 
`power.unkown.var`. Moreover, we specify a reasonable `max.n` which must be larger 
than `min.n`.

+ Step 3: Finally, by using the arguments from the step 2, we can compute the 
sample size for the situation with unknown covariance.
  
 
###### Step 1: 

* K= 2, number of co-primary endpoints 
* delta  = c(0.5, 0.4), expected effect size
* either Sigma or rho and SD must be given in the function
* Sigma = c(1, 0.5,0.5, 1), "known" covariance matrix
* rho = 0.5, "known" correlation
* SD = c(1,1), "known" standard deviation
* sig.level = 0.025, significance level 
* power = 0.80, intended power.

```{r}
Sigma <- matrix(c(1, 0.5, 0.5, 1), nrow = 2)
power.known.var(K = 2, delta = c(0.5, 0.4), Sigma = Sigma, sig.level = 0.025, power = 0.8)
```

###### Step 2:

The resulting value of `n`, which is 105, is used as `min.n`, and `max.n` must be larger then `min.n` so we try 120.

###### Step 3:
 
* K= 2, number of co-primary endpoints
* delta  = c(0.5, 0.4), expected effect size
* either Sigma or rho and SD must be given in the function
* Sigma = c(1, 0.5,0.5, 1), unknown covariance matrix
* rho = 0.5, unknown correlation 
* SD = c(1,1), unknown standard deviation
* sig.level = 0.025, significance level  
* power = 0.80,intended power
* min.n = 105, value of sample size with known variance
* max.n = 120, assuming a number which is greater then min.n
* n = ? , 

   
```{r}
Sigma <- matrix(c(1, 0.5, 0.5, 1), nrow = 2)
power.unknown.var(K = 2, delta = c(0.5, 0.4), Sigma = Sigma, sig.level = 0.025, power = 0.8, 
                  min.n = 105, max.n = 120)

```

```{r}
##equivalent: known SDs and correlation rho
Sigma <- matrix(c(1, 0.5, 0.5, 1), nrow = 2)
power.unknown.var(K = 2, delta = c(0.5, 0.4), rho = 0.5, SD = c(1,1), sig.level = 0.025, 
                  power = 0.8, min.n = 105, max.n = 120)

```

We obtain a sample size of n = 107 which is the sample size per group; that is, 
the total sample size required is 2n = 214. The results vary a little bit as
the calculation is based on `M` simulations (default: `M` = 10000).


Next, we calculate the power for a given sample size using the following setup.
   
* K= 2, number of co-primary endpoints 
* delta  = c(0.5, 0.4), expected effect size
* either Sigma or rho and SD must be given in the function
* Sigma = c(1, 0.5, 0.5, 1), expected covariance matrix
* rho = 0.5, expected correlation
* SD = c(1,1), expected standard deviations
* sig.level = 0.025, significance level  
* n = 107, sample size per group
* power = ?, calcuated power


```{r}
power.unknown.var(K = 2, n = 107, delta = c(0.5,0.4), Sigma = matrix(c(1, 0.5,0.5, 1), nrow=2), 
                  sig.level = 0.025)

```
The computed `power` is about $80.3\%$.
     

Moreover, we perform the corresponding intersection-union t-test by using simulated 
data and applying our function `mpe.t.test`


```{r}
## effect size
delta <- c(0.5, 0.4)
## covariance matrix
Sigma <- matrix(c(1, 0.5, 0.5, 1), ncol = 2)
## sample size
n <- 106
## generate random data from multivariate normal distributions
X <- rmvnorm(n=n, mean = delta, sigma = Sigma)
Y <- rmvnorm(n=n, mean = rep(0, length(delta)), sigma = Sigma)
## perform multivariate t-test
mpe.t.test(X = X, Y = Y)
```



## 3. Multiple Primary Endpoints
In this case, the trial is designed to find a significant difference for at least 
one endpoint. For example, in congestive heart failure trials like VEST [6] and  
MERIT [7] two primary endpoints are considered [5]. The primary endpoints are

1. All-cause mortality
2. Hospitalization for heart failure 

By increasing the number of endpoints the type I error increases. Hence, in 
order to control the type I error, one can apply the conservative Bonferroni 
correction [1, 10]. Here, we distribute $\alpha$ equally over all endpoints; 
i.e., $\alpha_k = \alpha/K (k = 1,...K)$. 

As we are considering a superiority trial, the superiority of the test intervention 
over control is given if and only if $µ_{T_k} - µ_{C_k} > 0$ for at least one 
$k = 1,2.....K$. In statistical testing, the null hypothesis is rejected if and 
only if at least one of the $K$ primary endpoints is significant at $\alpha_k$

   
   >$H_0 : µ_{T_k} - µ_{C_k} <= 0$, for all $k$,

   >$H_1 : µ_{T_k} - µ_{C_k} > 0$, for at least one $k$.
     
As we are assuming a known covariance, the hypotheses can be tested using a 
multivariate union-intersection z-test.


The overall power reads

$$1-\beta = P\left(\bigcup_{k=1}^K \{Z_k > z_\frac\alpha K\}\,\Big|\, H_1\right)$$

The overall power is also known as disjunctive power [3, 10].

We now demonstrate how to calculate the sample size for a trial with two 
multiple primary endpoints with known covariance.

* K = 2, number of primary endpoints
* delta = c(0.20,0.30), expected effect size 
* either Sigma or rho and SD must be given in the function
* Sigma = c(1, 0.3,0.3, 1), known covariance matrix
* rho = 0.3, correlation
* SD = c(1,1), known standard deviation
* power = 0.8, intended power
* n = ?


```{r}
atleast.one.endpoint(K = 2, delta = c(0.20,0.30),  Sigma = matrix(c(1, 0.3, 0.3, 1), nrow = 2),
                     power = 0.8, sig.level = 0.025)

```
```{r}
##equivalent: known SDs and correlation rho
atleast.one.endpoint(K = 2, delta = c(0.20, 0.30), SD = c(1,1), rho = 0.3,  power = 0.8)

```

   
We obtain a sample size of `n` = 147 which is the sample size per group; that is, the total sample size required is 2n = 294.

Later, we calculate the power for a given sample size using the following setup.
 
* K = 2, number of primary endpoints
* delta = c(0.20,0.30), expected effect size
* either Sigma or rho and SD must be given in the function
* Sigma = c(1, 0.3,0.3, 1), known covariance matrix
* rho = 0.3, known correlation
* SD = c(1,1), known standard deviation
* n = 147, sample size per group
* power = ?


```{r}
atleast.one.endpoint(K = 2, n = 147, delta = c(0.20,0.30),  
                     Sigma = matrix(c(1, 0.3, 0.3, 1), nrow = 2))

```
The computed `power` is $80.1\%$. 

## 4. Applications

###4.1. Multiple co-primary endpoints with known covariance 

In Alzheimer's Disease (AD) clinical trials, ADAS-cog (Alzheimer's Disease Assessment 
Scale - cognitive Sub-Scale) and CIBIC-plus (clinicians Interview Based Impression 
of change, Plus care give) are used as co-primary endpoints. In a 24 week double 
blinded, placebo controlled trial of donepezil in patients with AD [9], in which 
the standardized effect size of ADAS-cog and CIBIC-plus are given as 0.47 and 0.48 
and known correlations of the two endpoints are $\rho_z$ = 0.0, 0.3, 0.5, 0.8 at 
a significance level of 0.025 with `power = 0.80`, the sample sizes can be 
calculated as follows

* K = 2, number of co-primary endpoints 
* delta = c(0.47, 0.48), expected effect size
* either Sigma or rho and SD must be specified
* Sigma = matrix(c(1, rho, rho, 1), nrow = 2), known covariance matrix
* where rho = 0.0, or 0.3, or 0.5, or 0.8, known correlation 
* SD = c(1, 1), known standard deviations
* sig.level = 0.025, significance level 
* power = 0.80, intended power
* n = ?


```{r}
## correlation = 0.0,0.3,0.5,0.8
power.known.var(K = 2,delta = c(0.47, 0.48), Sigma = matrix(c(1, 0.0, 0.0, 1), nrow = 2), 
                sig.level = 0.025, power = 0.8)

power.known.var(K= 2,delta = c(0.47, 0.48), Sigma = matrix(c(1, 0.3, 0.3, 1), nrow = 2), 
                sig.level = 0.025, power = 0.8)

power.known.var(K = 2,delta = c(0.47, 0.48), Sigma = matrix(c(1, 0.5, 0.5, 1), nrow = 2), 
                sig.level = 0.025, power = 0.8)

power.known.var(K = 2, delta = c(0.47, 0.48), Sigma = matrix(c(1, 0.8, 0.8, 1), nrow = 2), 
                sig.level = 0.025, power = 0.8)

``` 
The sample sizes obtained at $\rho_z$ = 0.0, 0.3, 0.5, 0.8  are `n` = 92, 90, 87, 82 
for each group. That is, total sample sizes required are 
`2n`= 184, 180, 174, 164 and we observe that the sample size decreases with
increasing correlation.

###4.2. Multiple co-primary endpoints with unknown covariance    

We illustrate this by considering again a AD's clinical trial with the three 
co-primary endpoints SIB-score (an objective cognitive test), ADCS-ADL severity 
score (self-care and activities of daily living) and the CGI score (Clinical 
Global Impressions). The standardized effect size changes from baseline to six 
months are assumed as 0.36, 0.30, 0.26. The unknown correlations among the 
three endpoints are assumed to be identical at $\rho_z$ = 0.3. For 
`sig.level = 0.025` and `power = 0.80`, the sample size can be calculated as 
follows

###### step 1: 
Firstly, we assume the covariance as known and compute the sample size

* K = 3, number of co-primary endpoints
* delta = c(0.36, 0.30, 0.26), expected effect sizes 
* either Sigma or rho and SD must be given in the function
* Sigma = matrix(c(1, 0.3, 0.3, 0.3, 1, 0.3, 0.3, 0.3, 1), nrow = 3), "known" covariance matrix
* rho = c(0.3, 0.3, 0.3), "known" correlations
* SD = c(1, 1, 1), "known" standard deviations
* power = 0.8, intended power
* n = ?
   
```{r}

power.known.var(K = 3, delta = c(0.36, 0.30, 0.26),
                Sigma = matrix(c(1, 0.3, 0.3, 0.3, 1, 0.3, 0.3, 0.3, 1), nrow = 3),
                sig.level = 0.025, power = 0.8)

```

###### Step 2:  
The value of `n`, which is 268, is used as `min.n` and `max.n` must be larger 
then `min.n` so we try 300.

###### step 3: 
Now, we can compute the sample size with unknown covariance

* K = 3, number of co-primary endpoints
* delta = c(0.36, 0.30, 0.26), expected effect size 
* either Sigma or rho and SD must be given in the function
* Sigma = matrix(c(1, 0.3, 0.3, 0.3, 1, 0.3, 0.3, 0.3, 1), nrow = 3), expected covariance matrix
* rho = c(0.3, 0.3, 0.3), expected correlation
* SD = c(1, 1, 1), expected standard deviations
* power = 0.8, intended power
* min.n = 268, value of sample size with known variance
* max.n = 300, assuming a number which is greater then min.n
* n = ?
   
```{r}
power.unknown.var(K = 3, delta = c(0.36, 0.30, 0.26),
                Sigma = matrix(c(1, 0.3, 0.3, 0.3, 1, 0.3, 0.3, 0.3, 1), nrow = 3), 
                sig.level = 0.025, power = 0.8, min.n = 268, max.n = 300)

```
We obtain a sample size of `n` = 270 which is the sample size per group; that is, 
the total sample size required is `2n` = 540.
 
###4.3. Multiple primary endpoints

Here, by considering the example of the Alzheimer's Disease (AD) clinical trial 
discussed in Sec. 4.1, we illustrate the sample size calculations for 
multiple primary endpoints. The standardized effect sizes for ADAS-cog and 
CIBIC-plus are assumed as 0.47 and 0.48 and the correlation of the two endpoints 
is assumed to be known as $\rho_z$ = 0.0, 0.3, 0.8 at a significance level of 
0.025 with `power = 0.80`.

* K = 2, number of primary endpoints 
* delta = c(0.47, 0.48), expected effect size
* either Sigma or rho and SD must be specified
* Sigma = matrix(c(1, rho, rho, 1), nrow = 2), known covariance matrix
* where rho = 0.0, or 0.3, or 0.5, or 0.8, known correlation 
* SD = c(1, 1), known standard deviations
* sig.level = 0.025, significance level 
* power = 0.80, intended power
* n = ?

```{r}
atleast.one.endpoint(K= 2, delta = c(0.47, 0.48), Sigma = matrix(c(1, 0.0, 0.0, 1), nrow = 2), 
                     power = 0.8)

atleast.one.endpoint(K= 2, delta = c(0.47, 0.48), Sigma = matrix(c(1, 0.3, 0.3, 1), nrow = 2), 
                     power = 0.8)

atleast.one.endpoint(K= 2, delta = c(0.47, 0.48), Sigma = matrix(c(1, 0.5, 0.5, 1), nrow = 2),  
                     power = 0.8)

atleast.one.endpoint(K= 2, delta = c(0.47, 0.48), Sigma = matrix(c(1, 0.8, 0.8, 1), nrow = 2), 
                     power = 0.8)

```

Sample sizes obtained at $\rho_z$ = 0.0, 0.3, 0.5, 0.8 are `n` = 39, 45, 49, 57 
for each group. That is, the total sample sizes required are `2n`= 78, 90, 98, 104.
Hence, the sample size increases with increasing correlation.


##5. Summary

In clinical trials, the use of more then two primary endpoints is often necessary 
nowadays. Here, the correlation plays a crucial role which creates a quite
complicated situation and makes it challenging to calculate the sample size 
and to estimate power. Hence, in order to make this accessible, we follow the 
book [12] and implemented our package **mpe** with the functions introduced above. 
One can use it to calculate sample size and power for multiple (co-)primary 
endpoints and we hope that it will be useful for statisticians and practicians.
 

##6. References

1. Alex Dmitrenko, Ajit C. Tamhane, Frank Bret, _Multiple Testing Problems In Pharmaceutical statistics_. Chapman & Hall/CRC Biostatistics Series ISBN 9781584889847

2. Berger, R.L. (1982). _Multiparameter hypothesis testing and acceptance sampling_. Technometrics 24, 295-300

3. F.Bretz, T. Hothorn, and P. Westfall. _Multiple Comparisons Using R_. 2010 by Chapman and Hall/CRC ISBN 9781584885740 

4. Food and Drug Administration (FDA 2012). _Guidance for Industry Irritable Bowel Syndrome — Clinical Evaluation of Drugs for Treatment_. http://www.fda.gov/downloads/Drugs/.../Guidances/UCM205269.pdf

5. Gong J, Pinheiro JC, DeMets DL, _Estimating significance level and power comparisons for testing multiple endpoints in clinical trials.Control Clin Trials_. Control Clin Trials. 2000 Aug;21(4):313-29

6. J.N. Cohn, S.O. Goldstein, B.H. Greenberg, et al, _A dose-dependent increase in mortality with vesnarinone among patients with severe heart failure_. N Engl J Med, 339 (1998), pp. 1810–1816

7. MERIT-HF Study Group, _Effect of metoprolol CR/XL in chronic heart failure: Metoprolol CR/XL randomized intervention trial in congestive heart failure_. Lancet, 353 (1999), pp. 2001–2007

8. N. L. Johnson and S. Kotz (1972). _Distributions in Statistics—Continuous Multivariate Distributions,_ New York: John Wiley & Son

9. Rogers SL, Farlow MR, Doody RS, Mohs R, Friedhoff LT. _A 24-week, double-blind, placebo-controlled trial of donepezil in patients with Alzheimer's disease. Donepezil Study Group_. Neurology. 1998 Jan;50(1):136-45

10. Senn S, Bretz F, _Power and sample size when multiple endpoints are considered_. Pharm Stat. 2007 Jul-Sep;6(3):161-70

11. Sozu T, Sugimoto T, Hamasaki T (2011), _Sample Size Determination in Superiority Clinical Trials With Multiple Co-Primary Correlated Endpoints_. J Biopharm Stat. 2011 Jul;21(4):650-68

12. Sozu T, Sugimoto T, Hamasaki T, Evans S.R., _Sample Size Determination in Clinical Trials with Multiple Endpoints_. ISBN 978-3-319-22005-5


